Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-26 paper/main.tex
rename toc
write initial version of overview

\pagebreak
\renewcommand{\contentsname}{Temporary! Table of Contents}
\tableofcontents
\pagebreak

@@

%###################################################################################################
\section{Overview (TODO)}

\begin{itemize}
    \item This publication contains 5 things:
    \begin{enumerate}
        \item Methods of how to obtain data from microscopic images and perform parameter
            estimations; $\rightarrow$ also possibly across multiple generations
        \item Showcase the usefulness of the methods and the model by applying it to parameter
            estimation via microscopic images and multilayer formation
        \item Initial (incomplete) model of rod-shaped bacteria
        \item Generate 3D Images (masks)
        \item A python package which can do all of that and can be reused and extended
    \end{enumerate}
    \item Overall structure
    \begin{enumerate}
        \item users bring model + data
        \item this paper provides methods to do the parameter estimation + visualization
    \end{enumerate}
    \item For complete model, more work has to be done
    \begin{itemize}
        \item More aspects (see Review; put in discussion)
        \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
            important.
    \end{itemize}
    \begin{itemize}
        \item State of the Art
    \end{itemize}
\end{itemize}

\begin{figure}[H]
    \centering
    % \includegraphics[width=0.8\textwidth]{overview.pdf}
    \includestandalone[width=0.8\textwidth]{figures/concept/overview}
    \caption{
        The project can be divided into 4 subparts (yellow) that interact with and depend on each
        other.
        Additional blocks (blue-gray) indicate possible future extensions of this project.
        Arrows with solid lines indicate dependencies between the subparts while dashed arrows carry
        special meaning, which is specifically annotated.
        The mathematical models provides the theoretical framework and is thus the parent of all
        following work.
        "Data Extraction" allows us to read variable and parameter values directly from microscopic
        images.
        The obtained data can then be used to initialize a numerical simulations and thus estimate
        parameters.
        On the other hand, results produced by our model can be visualized as 3D renders and
        projected to cell-masks.
        These algorithms can be exploited to construct cost-functions which can be used in the
        parameter estimation process.
        In future work, this visualization scheme could be extended to generate realistic
        microscopic images with cell-masks and thus train cell-segmentation and cell-tracking
        algorithms.
    }
    \label{fig:flowchart-project-structure}
\end{figure}

>>>>

\pagebreak
\renewcommand{\contentsname}{Table of Contents (remove before submission)}
\tableofcontents
\pagebreak

@@

%###################################################################################################
\section{Overview (TODO)}
This work consists of many subcomponents which can be combined in order to achieve various tasks.
Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
This work assumes that bacterial rods can be represented by a colleciton of vertices $\vec{v}_i$.
This assumption alone is enough to allow us to visualize the bacteria and extract positional data
of bacterial agents from microscopic images.
The dynamics of the vertices $\vec{v}_i$ are determined by a computational model which is defined
by the respective researcher.
It is important to note that the components visualization and data extraction do not require any
knowledge about the computational model, thus allowing us to reuse their functionality across
various models.

Constructing a computational model requires knowledge about the target system and thus allows us to
interpret the results gathered by the provided data extraction methods which can then further be
used to restrict or determine the paramters of our model.
For example, by extracing the positions of bacteria across a time-series of microscopic images, we
can also calculate their lengths at these time-points and thus determine their rate of growth and
type of growth (see section~\ref{subsec:parameter-estimation-individual-treatment}).

The visualization methods are an integral part of any modeling endevaour.
They allow researchers to obtain feedback quickly and improve iteration times when constructing new
models.
Furthermore, they are an integral part of our advanced parameter estimation techniques
(see section~\ref{subsec:comparison-across-generations}) and enable us to effectively compare states
across generations of cells.
In the future, we are planning to extend this work and generate realistic microscopic images and
time-series.
Together with the parameter estimation methods, they can then be used to provide large amounts of
mechanistic data for the training cell-segmentation and cell-tracking machine learning algorithms.

% \begin{itemize}
%     \item This publication contains 5 things:
%     \begin{enumerate}
%         \item Methods of how to obtain data from microscopic images and perform parameter
%             estimations; $\rightarrow$ also possibly across multiple generations
%         \item Showcase the usefulness of the methods and the model by applying it to parameter
%             estimation via microscopic images and multilayer formation
%         \item Initial (incomplete) model of rod-shaped bacteria
%         \item Generate 3D Images (masks)
%         \item A python package which can do all of that and can be reused and extended
%     \end{enumerate}
%     \item Overall structure
%     \begin{enumerate}
%         \item users bring model + data
%         \item this paper provides methods to do the parameter estimation + visualization
%     \end{enumerate}
%     \item For complete model, more work has to be done
%     \begin{itemize}
%         \item More aspects (see Review; put in discussion)
%         \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
%             important.
%     \end{itemize}
%     \begin{itemize}
%         \item State of the Art
%     \end{itemize}
% \end{itemize}

\begin{figure}[H]
    \centering
    % \includegraphics[width=0.8\textwidth]{overview.pdf}
    \includestandalone[width=0.8\textwidth]{figures/concept/overview}
    \caption{
        The project can be divided into subcomponents (yellow, red) that interact with and depend on
        each other.
        Yellow blocks are provided by this work and red blocks require additional input such as
        modeling assumptions or input data.
        Blue blocks indicate possible future extensions of this project.
        Every component builds on the spatial representation of rod-shaped bacteria as vertices
        $\vec{v}_i$ for all proceeding work.
    }
    \label{fig:flowchart-project-structure}
\end{figure}
